Early Observations of the Type Ia Supernova \abc:\\ XXX NEED SOMETHING CLEVER TO SAY HERE XXX

### 2017-08-07 doc/ms.tex
Include comments from Jesper in the manuscript

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expansion and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
radius of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of
this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the explosion came from a 
star with $R_\ast \lesssim 0.02\sr$, where $\sr$ is the solar
radius. Combining the radius constraint with the measured ejecta 
mass \citeauthor{2012ApJ...744L..17B} derive the mean density of the 
progenitor star, confirming that at least some type Ia SNe come from 
compact and degenerate stars.

Early phase observations of SNe Ia from a white dwarf (WD)$+$non-degenerate binary
may detect excess emission, relative to most type Ia SNe, due to the 
collision of the SN ejecta with the non-degenerate companion 
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess 
emission was first detected in iPTF\,14atg  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as a
SN ejecta-companion collision (but see also \citealt{2016MNRAS.459.4428K,2017arXiv170603613N}). While such emission requires a 
favorable geometric alignment and is only expected in $\lesssim$ 
10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have 
searched for signatures of an ejecta-companion interaction, 
typically resulting in non-detections 
(e.g., \citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}). Possible exceptions include SN\,2012cg, 
which exhibited excess blue emission in its early-phase light curve 
(\citealt{2016ApJ...820...92M}; though see 
\citealt{2016arXiv161007601S} for an interpretation that does not 
invoke ejecta-companion interaction), and SN\,2017cbv, which shows a clearly resolved ``bump'' in the early $UBg$ light curves \citep{2017arXiv170608990H}. 
% With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
% signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

@@

The vast majority of SNe Ia are observed 
to be powered purely by the radioactive decay of $^{56}$Ni. 
While the detection of SN shock cooling or ejecta interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can fundamentally alter the appearance of the SN shortly after explosion. SNe Ia experience a dark period after
the SN shock breakout but before radioactive energy  
diffuses into the photosphere \citep{2014ApJ...784...85P}. The 
duration of this dark period is set by how the newly synthesized 
$^{56}$Ni is mixed and deposited into different layers of the ejecta. In the case of strong mixing, the dark period is short, or 
non-existent, as photons from $^{56}$Ni decay can reach the 
photosphere rapidly. This also leads to blue optical colors and a 
rapid initial rise in the observed light curve. If the mixing is 
weak and the $^{56}$Ni is confined to the innermost layers of the 
ejecta, the dark period can last for several days as the radioactive energy diffuses to the photosphere. The early evolution of such SNe results in redder colors and a more moderate rise in luminosity \citep{2016ApJ...826...96P}. Thus, the early light curves of even non-exotic type Ia SNe convey information about their progenitor systems by constraining the distribution of synthesized $^{56}$Ni, which in turn constrains the explosion mechanism.

In this paper, we report observations of an extraordinarily young SN Ia, \abc, which was discovered by the intermediate Palomar Transient Factory (iPTF) on 2016 April $3.36$ UTC at $\textrm{R.A.}=13^h34^m45.49^s$, $\textrm{Dec.}=+13^d51^m14.3^s$ (J2000) with a $g_\mathrm{PTF}$-band magnitude of $21.31\pm0.27$ \citep{2016ATel.8907....1M}. The transient is spatially coincident with a tidal tail of the galaxy NGC\,5221, which lies at a distance of $\sim$100\,Mpc. \abc\ is not detected to a limit of $g=22.1$\,mag on April $2.42$, less than 1 d prior to discovery, and rose by $\sim$2 mag in the 24 hr following its initial detection. Our spectroscopic follow-up
campaign classified \abc\ as a normal SN Ia
\citep{2016ATel.8909....1C}. Our observations and analysis show that the early evolution of \abc\ exhibited several distinct properties relative to SN\,2011fe. We interpret those differences as arising from either strong $^{56}$Ni mixing or ejecta interaction with diffuse material, or a combination of the two.

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=0.48\textwidth]{logLC_with_colors.pdf}
  \caption{The $g$-band
  light curve of \abc. Observations from different 
  telescopes are shown with different symbols. Time is measured in 
  rest-frame d relative to the time of first light, $t_0$ (see 
  \S\ref{sec:lc_fit}). Note that the horizontal axis is shown with 
  a linear scale for $-2 \; \mathrm{d} \le t - t_0 \le 3 \; 
  \mathrm{d}$ and a $\log$ scale for $t - t_0 > 3 \; \mathrm{d}$. 
  The black ticks near the
  bottom of the panel show epochs of spectroscopic observations.
  }
  \label{fig:lightcurve}
\end{figure}

@@

After discovery, photometric observations in the $g'$, $r'$ and $i'$
filters were obtained with the SED Machine 
(SEDm; Blagorodnova et al.\ 2017, in prep.) mounted on the Palomar Observatory 
60-inch telescope (P60). We utilized the Fremling Automated Pipeline \citep{2016A&A...593A..68F} to subtract galaxy light from the SEDm images using archival Sloan Digital Sky Survey (SDSS) images as a reference. This pipeline then performed forced-PSF photometry at the location of \abc, which is calibrated to the SDSS catalog \citep{2014ApJS..211...17A}.

Photometric observations in the $BVg'r'i'$ filters were conducted by the Las Cambres Observatory (LCO) 1-m
telescope network.  PSF photometry was measured on these images using
the \texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The
\textit{BV} magnitudes are calibrated to the Fourth USNO CCD
Astrograph Catalog \citep{2013AJ....145...44Z}, and the \textit{g'r'i'}
magnitudes are calibrated to SDSS Data Release 6
\citep{2008ApJS..175..297A}.

@@

Spectroscopic observations of \abc\ were taken with a variety
of telescopes and instruments over multiple epochs spanning from a
couple of days after explosion to two months after $B$-band maximum. 
An observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were reduced using standard routines in \texttt{IDL}/\texttt{Python}. The optical
spectral evolution of \abc\ is illustrated in Figure
\ref{fig:spec_seq}, which excludes high-resolution Very Large Telescope (VLT) spectra for clarity.

@@

A detailed study of the reddening towards \abc\ is presented in a 
companion paper (Ferretti et al.\ 2017, submitted). Briefly, the 
foreground Galactic extinction toward \abc\
is $E(B-V) = 0.0279 \, \mathrm{mag}$ \citep{2011ApJ...737..103S}. The high-resolution spectra presented in Ferretti et al.\ (2017) show multiple absorption components for both the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets. While the equivalent width (EW) of these lines is quite large, implying a large amount of extinction (e.g., \citealt{2012MNRAS.426.1465P}), Ferretti et al.\ compare the evolution of \abc\ to the well-observed normal type Ia SN\,2011fe and find evidence for only a small amount of extinction. The empirical relation between the EW of \ion{Na}{1}\,D and extinction is known to have a large scatter, and \citet{2013ApJ...779...38P} have shown that \ion{Na}{1}\,D absorption is a poor tracer of reddening in Type Ia SNe. Thus, we instead adopt $E(B-V) = 0.05 \, \mathrm{mag}$ as the local extinction for \abc\ (Ferretti et al.\ 2017). For the remainder of our analysis we therefore adopt a total, Galactic $+$ host galaxy, line-of-sight extinction of $E(B-V) = 0.08 \, \mathrm{mag}$.

@@

To determine the brightness and time of $B$-band maximum for 
\abc, we fit the P60 light curves with the \texttt{sncosmo} software package.\footnote{The
  \texttt{sncosmo} Python module is available at
  \url{https://sncosmo.readthedocs.io/en/v1.5.x/}.} This fit includes a \texttt{SALT2} template \citep{2007A&A...466...11G} that has been corrected for extinction using the \citep{1999PASP..111...63F} reddening law, $R_V=3.1$, and $E(B-V)$ values from \S\ref{sec:reddening}.

@@

The time of first light for SNe is usually estimated by 
extrapolating early-phase light curves to determine when the 
SN is equal to 0. Assuming an ideal expanding fireball with 
constant temperature, \citet{1982ApJ...253..785A} derives that $f 
\propto t^2$, where $f$ is the SN flux and $t$ is the time since 
explosion. Despite the basic assumption of a constant temperature at 
early times, multiple studies have found that the early emission 
from type Ia SNe can be described as a power law in time, with 
power-law index consistent with 2, i.e. $f \propto t^2$ (e.g., \citealt{2006AJ....132.1707C, 2010ApJ...712..350H, 2011MNRAS.416.2607G}).\footnote{Many of the studies that find $f \propto t^2$ sample SNe Ia light curves at a phase that is closer to $T_{B,\mathrm{max}}$ than the initial observations of \abc\ discussed here.}

As our observations include especially early observations of \abc, there are upper limits $\sim$1 d prior to the discovery epoch, we model the early flux from \abc\ as a power law, but allow the power-law index to vary, as opposed to fixing it at 2, to account for potential variations in the photospheric temperature during expansion:
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t<t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ ,
\end{equation}
where $t_0$ is the time of first light, $\alpha$ is the power-law index, and $t$ is measured in the rest-frame of the SN. To determine $t_0$ and $\alpha$ we fit the earliest observations of \abc. Due to slight variations in the passbands, we fit the model only to the relative flux measured in the $g_\mathrm{PTF}$-band, which is the only filter with observations prior to first light, a necessity for constraining $t_0$.

To determine the best fit parameters, we search a large grid over $t_0$, $\alpha$
and the proportionality constant, and minimize $\chi^2$. 
The modeling results show that the SN flux rises approximately
linearly between
$t=-18\,\textrm{d}$ and $t=-14\,\textrm{d}$. Figure \ref{fig:early_lc_fit} shows the best-fit result and
the joint marginal distribution of $t_0$ and $\alpha$. From the
best-fit model we obtain $\alpha=0.98 \pm ^{0.16}_{0.14}$ and $t_0=-17.91 \pm ^{0.07}_{0.15}$ d, where the uncertainties represent the marginalized 95\% confidence intervals. Our first
detection of \abc\ occurred $\sim{0.15}\,\textrm{d}$ after 
the SN first light. Figure~\ref{fig:early_lc_fit} additionally shows $g'$-bands observations from P60 and LCO, where we have normalized the flux assuming that the $g$ magnitudes from each instrument have the same zero-point. This assumption is incorrect (see \S\ref{sec:obs}), and, as expected, the residuals show a systematic offset between the $g_\mathrm{PTF}$-band and the $g'$-band. Nevertheless, if we ignore this systematic and fit all of the $g$-band observations simultaneous we obtain marginalized best-fit parameters that are consistent within the uncertainties with the $g_\mathrm{PTF}$-only values above. In the analysis that follows, the precise values of the best-fit parameters is not important. The critical finding here is that $\alpha \approx 1$ and $t_0 \approx -18 \, \mathrm{d}$.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.48\textwidth]{early_lc.pdf}
  \caption{Best-fit power-law model to describe the early flux from 
  \abc\ in the $g_\mathrm{PTF}$-band. The model flux, adopting 
  best-fit parameters $\alpha=0.98$ and $t_0=-17.94\,\textrm{d}$, 
  is shown as a solid black line in the top panel. The lower panel 
  shows the model residuals, and the symbols are the same as in
  Figure~\ref{fig:lightcurve}. The model is fit to just the 
  $g_\mathrm{PTF}$-band observations to avoid systematic difference 
  between the filters (see the text for further 
  details). The joint distribution of $t_0$ and $\alpha$ is 
  illustrated in the inset of the upper panel. The solid and 
  dashed contours represent the $68\%$ and $99.7\%$ confidence 
  levels.
  }
  \label{fig:early_lc_fit}
\end{figure}

% Starting around $t=-15\,\textrm{d}$, the $g$-band light curve
% deviates from the above best-fit model and
% rises significantly faster, indicating a
% change in the power-law index $\alpha$. In fact, between $t=-14\,\mathrm{d}$ and $t=-8\,\mathrm{d}$ the light curve is well fit by a power law of index $1.40$. In other words, the entire light curve
% before $t=-8$\,days can be approximated by a broken power-law model
% (e.g.,
% \citealt{2016arXiv161202097Z,
%   2016arXiv161202725Z}).

As previously noted, the early emission from most SNe Ia can be explained via a $f \propto t^2$ model, including SNe with extremely early observations like \abc, such as SN\,2011fe \citep{2011Natur.480..344N}. Thus, the near-linear rise in flux over the first few days after first light for \abc\ is unusual. To our knowledge this behavior has only been observed in 2 other SNe (2013dy, 2014J; \citealt{2013ApJ...778L..15Z,2014ApJ...783L..24Z,2015ApJ...799..106G}). Any model to explain the observations of \abc\ must be able to account for this unusual behavior in the days after first light.

@@

As previously noted, $t_0$ does not correspond to the time of 
explosion, $t_\mathrm{exp}$, for type Ia SNe as the SN may 
experience a dark phase following shock breakout before radioactive 
energy can diffuse to the photosphere. \citet{2014ApJ...784...85P} 
instead suggest that measurements of the photospheric velocity be 
used to determine $t_\mathrm{exp}$ given that the ejecta begin 
expanding from the moment of explosion.  Assuming a
constant opacity in the ejecta, \citeauthor{2014ApJ...784...85P} 
find that the photospheric velocity evolves as
$v_{ph}\propto(t-t_\mathrm{exp})^{-0.22}$. Numerical experiments by \citet{2016ApJ...826...96P} find that the consatnt-opacity assumption strongly depends on the amount of $^{56}$Ni mixing in the SN ejecta (for a more detailed discussion of $^{56}$Ni mixing see \S\ref{sec:Ni_mixing}). As a result, the adoption of a $t^{-0.22}$ power-law model may not be valid for all SNe Ia. Nevertheless, we proceed on the assumption that \abc\ experienced strong $^{56}$Ni mixing, corresponding to the models that are best approximated as a $t^{-0.22}$ power law. We do this in part to compare with previous studies, though we caution that our derived value for $t_{exp}$ is subject to uncertainties related to ejecta mixing.

While the photospheric velocity
is not easy to measure, line velocities of \ion{Si}{2} or \ion{Ca}{2}
can be used as a proxy
\citep{2014ApJ...784...85P,2016ApJ...826..144S}. In the case of \abc, the \ion{Ca}{2} IR triplet is very weak, 
likely due to high temperatures in the ejecta. Thus, we
determine the photospheric velocity from the 
\ion{Si}{2}\,$\lambda$6355 line. Visual inspection shows no 
sign of multi-velocity components of \ion{Si}{2}, and that 
the \ion{C}{2}\,$\lambda$6580 line overlaps the red
wing of the \ion{Si}{2} line (see Figures~\ref{fig:spec_seq} and \ref{fig:carbon}). Consequently, we model the observed 
spectra between 5,900 and 6,500\,\AA\ (rest-frame) as the combination 
of two gaussian kernels plus a linear baseline, which
accounts for \ion{Si}{2}, \ion{C}{2} and the
continuum, respectively. The expansion
velocity of \ion{Si}{2} is measured by the central wavelength
of the \ion{Si}{2} Gaussian kernel.

>>>>

While the shock breakout of a SN Ia occurs on a sub-second timescale,
the subsequent quasi-adiabatic expansion and cooling of the unbound
ejecta produces thermal emission that can be used to infer the
radius of the exploding star
\citep{2010ApJ...708..598P,2011ApJ...728...63R}. Comparing models of this cooling emission to the earliest-phase data of SN~2011fe,
\citet{2012ApJ...744L..17B} concluded that the explosion came from a star with $R_\ast \lesssim 0.02\sr$, where $\sr$ is the solar
radius. Combining the radius constraint with the measured ejecta 
mass, \citeauthor{2012ApJ...744L..17B} derive the mean density of the progenitor star, confirming that at least some Type Ia SNe come from compact and degenerate stars.

Early phase observations of SNe Ia from a white dwarf (WD)$+$non-degenerate binary
may detect excess emission, relative to most Type Ia SNe, due to the 
collision of the SN ejecta with the non-degenerate companion 
\citep{1973ApJ...186.1007W,2010ApJ...708.1025K}. This excess 
emission was first detected in iPTF\,14atg  \citep{2015Natur.521..328C}, a low-velocity SN Ia with a significant and declining ultraviolet (UV) pulse detected within a
few days of the SN explosion. This UV pulse is best interpreted as a
SN ejecta-companion collision (but see also \citealt{2016MNRAS.459.4428K,2017arXiv170603613N}). While such emission requires a 
favorable geometric alignment and is only expected in $\lesssim$ 
10\% of SNe Ia \citep{2010ApJ...708.1025K}, many studies have 
searched for signatures of an ejecta-companion interaction, 
typically resulting in non-detections 
(e.g., \citealt{2010ApJ...722.1691H,2011ApJ...741...20B,2012ApJ...744...38F,
  2012ApJ...744L..17B,2015Natur.521..332O,
  2013ApJ...778L..15Z,2015ApJ...799..106G,2016ApJ...826..144S,
  2015ApJS..221...22I}). Possible exceptions include SN\,2012cg, 
which exhibited excess blue emission in its early-phase light curve 
(\citealt{2016ApJ...820...92M}; though see 
\citealt{2016arXiv161007601S} for an interpretation that does not 
invoke ejecta-companion interaction), and SN\,2017cbv, which shows a clearly resolved ``bump'' in the early $UBg$ light curves \citep{2017arXiv170608990H}. 
% With only $\sim 10\%$ of single-degenerate progenitors occupying the preferred binary geometry for us to see the collision
% signatures, it is not surprising that there are few confirmed detections in the literature \ycao{ref?}.

@@

The vast majority of SNe Ia are observed 
to be powered purely by the radioactive decay of $^{56}$Ni. 
While the detection of SN shock cooling or ejecta interaction is rare, the level of $^{56}$Ni mixing in the SN ejecta can fundamentally alter the appearance of the SN shortly after explosion. SNe Ia experience a dark period after
the SN shock breakout but before radioactive energy  
diffuses into the photosphere \citep{2014ApJ...784...85P}. The 
duration of this dark period is set by how the newly synthesized 
$^{56}$Ni is mixed and deposited into different layers of the ejecta. In the case of strong mixing, the dark period is short, or 
non-existent, as $\gamma$-rays from the $^{56}$Ni decay diffuse to the photosphere rapidly. This also leads to blue optical colors and a rapid initial rise in the observed light curve. If the mixing is weak and the $^{56}$Ni is confined to the innermost layers of the ejecta, the dark period can last for several days as the radioactive energy diffuses to the photosphere. The early evolution of such SNe results in redder colors and a more moderate rise in luminosity \citep{2016ApJ...826...96P}. Thus, the early light curves of even non-exotic Type Ia SNe convey information about their progenitor systems by constraining the distribution of synthesized $^{56}$Ni, which in turn constrains the explosion mechanism.\footnote{While examing a sample of Type Ib/c SNe, \citet{2015A&A...574A..60T} found that a significant dark phase is not prevalent for core-collapse SNe.}

\citet{2017arXiv170603613N} recently demonstrated that disambiguating between these different scenarios via optical photometry alone is challenging. \citeauthor{2017arXiv170603613N} further show that estimates of the time of explosion, which are critical for comparing models with observations, are often incorrect by as many as $\sim$2 d using common methods in the literature. While analytical models suggest that early spectra can be used to infer the time of explosion (e.g., \citealt{2014ApJ...784...85P}), more detailed simulations show that the photospheric evolution is not so simple \citep{2016ApJ...826...96P}. Reconciling these issues requires both a larger sample of early SNe Ia observations and more detailed models that produce synthetic light curves and spectra.

In this paper, we report observations of an extraordinarily young SN Ia, \abc, which was discovered by the intermediate Palomar Transient Factory (iPTF) on 2016 April $3.36$ UTC at $\textrm{R.A.}=13^h34^m45.49^s$, $\textrm{Dec.}=+13\degr51\arcmin14\farcs3$ (J2000) with a $g_\mathrm{PTF}$-band magnitude of $21.31\pm0.27$ \citep{2016ATel.8907....1M}. The transient is spatially coincident with a tidal tail of the galaxy NGC\,5221, which lies at a distance of $\sim$100\,Mpc. \abc\ is not detected to a limit of $g=22.1$\,mag on April $2.42$, less than 1 d prior to discovery, and rose by $\sim$2 mag in the 24 hr following its initial detection. Our spectroscopic follow-up
campaign classified \abc\ as a normal SN Ia
\citep{2016ATel.8909....1C}. Our observations and analysis show that the early evolution of \abc\ exhibited several distinct properties relative to SN\,2011fe. We interpret those differences as arising from either strong $^{56}$Ni mixing or ejecta interaction with diffuse material, or a combination of the two.

@@

\begin{figure}[htb]
  \centering
  \includegraphics[width=0.48\textwidth]{logLC_with_colors.pdf}
  \caption{The $g$-band
  light curve of \abc. Observations from different 
  telescopes are shown with different symbols. Time is measured in 
  rest-frame days relative to the time of first light, $t_0$ (see 
  \S\ref{sec:lc_fit}). Note that the horizontal axis is shown with 
  a linear scale for $-2 \; \mathrm{d} \le t - t_0 \le 3 \; 
  \mathrm{d}$ and a $\log$ scale for $t - t_0 > 3 \; \mathrm{d}$. 
  The black ticks near the
  bottom of the panel show epochs of spectroscopic observations.
  }
  \label{fig:lightcurve}
\end{figure}

@@

After discovery, photometric observations in the $g'$, $r'$ and $i'$
filters were obtained with the SED Machine 
(SEDm; Blagorodnova et al.\ 2017, in prep.) mounted on the Palomar Observatory 
60-inch telescope (P60). We utilized the Fremling Automated Pipeline (\texttt{FPipe}; \citealt{2016A&A...593A..68F}) to subtract galaxy light from the SEDm images using archival Sloan Digital Sky Survey (SDSS) images as a reference. This pipeline then performed forced-PSF photometry at the location of \abc, which is calibrated to the SDSS catalog \citep{2014ApJS..211...17A}.

Photometric observations in the $BVg'r'i'$ filters were conducted by the Las Cumbres Observatory (LCO) 1-m
telescope network.  PSF photometry was measured on these images using
the \texttt{lcogtsnpipe} pipeline \citep{2016MNRAS.459.3939V}. The
\textit{BV} magnitudes are calibrated to the Fourth USNO CCD
Astrograph Catalog \citep{2013AJ....145...44Z}, and the \textit{g'r'i'}
magnitudes are calibrated to SDSS Data Release 6
\citep{2008ApJS..175..297A}.

@@

Spectroscopic observations of \abc\ were taken with a variety
of telescopes and instruments over multiple epochs spanning from a
couple of days after explosion to two months after $B$-band maximum. 
An observing log is listed in Table \ref{tab:spec_obs_log}. The spectra were reduced using standard routines in \texttt{IDL}/\texttt{Python}/\texttt{Matlab}. The optical
spectral evolution of \abc\ is illustrated in Figure
\ref{fig:spec_seq}, which excludes high-resolution Very Large Telescope (VLT) spectra for clarity.

@@

A detailed study of the reddening towards \abc\ is presented in a 
companion paper (Ferretti et al.\ 2017, submitted). Briefly, the 
foreground Galactic extinction toward \abc\
is $E(B-V) = 0.0279 \, \mathrm{mag}$ \citep{2011ApJ...737..103S}. The high-resolution spectra presented in Ferretti et al.\ (2017) show multiple absorption components for both the \ion{Ca}{2}\,H$+$K and \ion{Na}{1}\,D doublets. While the equivalent width (EW) of these lines is quite large, implying a large amount of extinction (e.g., \citealt{2012MNRAS.426.1465P}), Ferretti et al.\ compare the evolution of \abc\ to the well-observed normal Type Ia SN\,2011fe and find evidence for only a small amount of extinction. The empirical relation between the EW of \ion{Na}{1}\,D and extinction is known to have a large scatter, and \citet{2013ApJ...779...38P} have shown that \ion{Na}{1}\,D absorption is a poor tracer of reddening in Type Ia SNe. Thus, we instead adopt $E(B-V) = 0.05 \, \mathrm{mag}$ as the local extinction for \abc\ (Ferretti et al.\ 2017). For the remainder of our analysis we therefore adopt a total, Galactic $+$ host galaxy, line-of-sight extinction of $E(B-V) = 0.08 \, \mathrm{mag}$.

@@

To determine the brightness and time of $B$-band maximum for 
\abc, we fit the P60 light curves with the \texttt{sncosmo} software package.\footnote{The
  \texttt{sncosmo} Python module is available at
  \url{https://sncosmo.readthedocs.io/en/v1.5.x/}.} This fit includes a \texttt{SALT2} template \citep{2007A&A...466...11G} that has been corrected for extinction using the \citep{1999PASP..111...63F} reddening law, $R_V=3.1$, and $E(B-V) = 0.08 \; \mathrm{mag}$.

@@

The time of first light for SNe is usually estimated by 
extrapolating early-phase light curves to determine when the 
SN flux is equal to 0. Assuming an ideal expanding fireball with 
constant temperature, \citet{1982ApJ...253..785A} derives that $f 
\propto t^2$, where $f$ is the SN flux and $t$ is the time since 
explosion. Despite the basic assumption of a constant temperature at 
early times, multiple studies have found that the early emission 
from Type Ia SNe can be described as a power law in time, with 
power-law index consistent with 2, i.e. $f \propto t^2$ (e.g., \citealt{2006AJ....132.1707C, 2010ApJ...712..350H, 2011MNRAS.416.2607G}).\footnote{Many of the studies that find $f \propto t^2$ sample SNe Ia light curves at a phase that is closer to $T_{B,\mathrm{max}}$ than the initial observations of \abc\ discussed here.}

As our observations include especially early observations of \abc, there are upper limits $\sim$1 d prior to the discovery epoch, we model the early flux from \abc\ as a power law, but allow the power-law index to vary, as opposed to fixing it at 2, to account for potential variations in the photospheric temperature during expansion:
\begin{equation}
  \label{eq:broken_power_law}
  f(t) \left\{
    \begin{array}{ll}
      = 0,\ \textrm{when}\ t \le t_0 \\
      \propto (t-t_0)^{\alpha},\ \textrm{when}\ t>t_0
    \end{array}
  \right.\ ,
\end{equation}
where $t_0$ is the time of first light, $\alpha$ is the power-law index, and $t$ is measured in the rest-frame of the SN. To determine $t_0$ and $\alpha$ we fit the earliest observations of \abc. Due to slight variations in the passbands, we fit the model only to the relative flux measured in the $g_\mathrm{PTF}$-band, which is the only filter with observations prior to first light, a necessity for constraining $t_0$.

To determine the best fit parameters, we search a large grid over $t_0$, $\alpha$
and the proportionality constant, and minimize $\chi^2$. 
The modeling results show that the SN flux rises approximately
linearly between
$t=-18\,\textrm{d}$ and $t=-14\,\textrm{d}$. Figure \ref{fig:early_lc_fit} shows the best-fit result and
the joint marginal distribution of $t_0$ and $\alpha$. From the
best-fit model we obtain $\alpha=0.98 \pm ^{0.16}_{0.14}$ and $t_0=-17.91 \pm ^{0.07}_{0.15}$ d, where the uncertainties represent the marginalized 95\% confidence intervals. Our first
detection of \abc\ occurred $\sim$${0.15}\,\textrm{d}$ after 
the SN first light. 
% Figure~\ref{fig:early_lc_fit} additionally shows $g'$-bands observations from P60 and LCO, where we have normalized the flux assuming that the $g$ magnitudes from each instrument have the same zero-point. This assumption is incorrect (see \S\ref{sec:obs}), and, as expected, the residuals show a systematic offset between the $g_\mathrm{PTF}$-band and the $g'$-band. Nevertheless, if we ignore this systematic and fit all of the $g$-band observations simultaneously we obtain marginalized best-fit parameters that are consistent within the uncertainties with the $g_\mathrm{PTF}$-only values above.
In the analysis that follows, the precise values of the best-fit parameters are not important. The critical finding here is that $\alpha \approx 1$ and $t_0 \approx -18 \, \mathrm{d}$.

As previously noted, a precise determination of the rise time, $t_\mathrm{rise}$, of SNe Ia is challenging as there may be a ``dark phase'' following explosion \citep{2014ApJ...784...85P}. Nevertheless, to be consistent with previous studies (e.g., \citealt{2011MNRAS.416.2607G}), here we find $t_rise = 17.91\pm _{0.07}^{0.15}$ d based on our fit for $t_0$. We caution that this measurement may be underestimated due to the uncertain time of explosion.

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.48\textwidth]{early_lc.pdf}
  \caption{Best-fit power-law model to describe the early flux from 
  \abc\ in the $g_\mathrm{PTF}$-band. The model flux, adopting 
  best-fit parameters $\alpha=0.98$ and $t_0=-17.94\,\textrm{d}$, 
  is shown as a solid black line in the top panel. The lower panel 
  shows the model residuals, and the symbols are the same as in
  Figure~\ref{fig:lightcurve}. The model is fit to just the 
  $g_\mathrm{PTF}$-band observations to avoid systematic difference 
  between the filters (see the text for further 
  details). The joint distribution of $t_0$ and $\alpha$ is 
  illustrated in the inset of the upper panel. The solid and 
  dashed contours represent the $68\%$ and $99.7\%$ confidence 
  levels. \amiller{Update figure to remove other photometry}
  }
  \label{fig:early_lc_fit}
\end{figure}

Simulations presented in \citet{2017arXiv170603613N} show that the flux from SNe Ia explosion models does not evolve as a power law in time. Despite this non-power-law evolution, \citet{2017arXiv170603613N} show that $f \propto t^\alpha$ models produce good formal fits to simulated observations. The $f \propto t^\alpha$ fits to the \citeauthor{2017arXiv170603613N} models result in glaring errors to the estimated explosion times. Despite the caution advised in \citet{2017arXiv170603613N}, we proceed with the power-law fits from above for several reasons. (1) Our primary aim with the fit is to characterize $\alpha$ for \abc\ compared to SN\,2011fe (see below). (2) The errors in the estimated explosion times from \citeauthor{2017arXiv170603613N} increase the later the first observation of the SN is obtained. \citeauthor{2017arXiv170603613N} only present fits to observations that start $>$2~d after explosion, whereas the discovery epoch for \abc\ is likely much earlier than that (see \S\ref{sec:early_vel}). (3) We do not use the fits to estimate the time of explosion, $t_0$ instead is an estimate of the time of first light. (4) Pre-discovery upper limits, which are not included in \citeauthor{2017arXiv170603613N}, provide additional strong constraints on the time of first light. Finally, (5) \citet{2017arXiv170603613N} find that models with strong $^{56}$Ni mixing do follow a power-law evolution, and we explore, and argue in favor of, this possibility for \abc\ (see \S\ref{sec:Ni_mixing}).

As previously noted, the early emission from SN\,2011fe is well fit by a $f \propto t^2$ model \citep{2011Natur.480..344N}. Thus, the near-linear rise in flux over the first few days after first light for \abc\ is distinct compared to SN\,2011fe. To our knowledge this behavior has only been observed in 2 other SNe (2013dy, 2014J; \citealt{2013ApJ...778L..15Z,2014ApJ...783L..24Z,2015ApJ...799..106G}). Any model to explain the observations of \abc\ must account for this near-linear rise in the days after first light.

@@

As previously noted, $t_0$ may not correspond to the time of 
explosion, $t_\mathrm{exp}$, for Type Ia SNe as the SN may 
experience a dark phase following shock breakout before radioactive 
energy can diffuse to the photosphere. \citet{2014ApJ...784...85P} 
instead suggest that measurements of the photospheric velocity be 
used to determine $t_\mathrm{exp}$ given that the ejecta begin 
expanding from the moment of explosion.  Assuming a
constant opacity in the ejecta, \citeauthor{2014ApJ...784...85P} 
find that the photospheric velocity evolves as
$v_{ph}\propto(t-t_\mathrm{exp})^{-0.22}$. Numerical experiments by \citet{2016ApJ...826...96P} find that the consatnt-opacity assumption strongly depends on the amount of $^{56}$Ni mixing in the SN ejecta (for a more detailed discussion of $^{56}$Ni mixing see \S\ref{sec:Ni_mixing}). As a result, the adoption of a $t^{-0.22}$ power-law model may not be valid for all SNe Ia. Nevertheless, we proceed on the assumption that \abc\ experienced strong $^{56}$Ni mixing, corresponding to the models that are best approximated as a $t^{-0.22}$ power law. We do this in part to compare with previous studies, though we caution that our derived value for $t_{exp}$ is subject to uncertainties related to ejecta mixing.

While the photospheric velocity
is not easy to measure, line velocities of \ion{Si}{2} or \ion{Ca}{2}
can be used as a proxy
\citep{2014ApJ...784...85P,2016ApJ...826..144S}. In the case of \abc, the \ion{Ca}{2} IR triplet is very weak, 
likely due to high temperatures in the ejecta. Thus, we
determine the photospheric velocity from the 
\ion{Si}{2}\,$\lambda$6355 line. Visual inspection shows no 
sign of multi-velocity components of \ion{Si}{2}, and that 
the \ion{C}{2}\,$\lambda$6580 line overlaps the red
wing of the \ion{Si}{2} line (see Figures~\ref{fig:spec_seq} and \ref{fig:carbon}). Consequently, we model the observed 
spectra between 5,900 and 6,500\,\AA\ (rest-frame) as the combination 
of two Gaussian kernels plus a linear baseline, which
accounts for \ion{Si}{2}, \ion{C}{2} and the
continuum, respectively. The expansion
velocity of \ion{Si}{2} is measured by the central wavelength
of the \ion{Si}{2} Gaussian kernel.

\begin{figure*}[!thb]
  \centering
  \includegraphics[width=0.48\textwidth]{Chi2.pdf}
  \includegraphics[width=0.48\textwidth]{VelocityPlot.pdf}
  \caption{Constraints on $t_{exp}$ from fitting the velocity
    evolution of \ion{Si}{2}.
    \textit{Left panel:} the dashed, solid and dash-dotted curves 
    show $\chi^2$ for fitting power laws with
    indices $-0.20$, $-0.22$ and $-0.24$, respectively. The blue
    vertical line and the orange shaded region indicate $t_0$ and 
    its 95\% confidence interval from Section
    \ref{sec:lc_fit}, respectively.
    \textit{Right panel:} Observed \ion{Si}{2}\,$\lambda$6355
    velocities (blue circles) and the best-fit power-law model
     with an index of $-0.22$ (dashed line). Additionally, the 
     measured velocites of \ion{C}{2}\,$\lambda\lambda$6580, 7234 
     are shown. Typical uncertainties are smaller than the size of 
     the points. \textit{Right inset}: Evolution of the 
     pseudo-equivalent width of \ion{Si}{2}\,$\lambda$6355, and 
     \ion{C}{2}\,$\lambda\lambda$6580, 7234 in the week following 
     explosion.}
  \label{fig:velocity_t_exp}
\end{figure*}
